---
title: "DataCleaning"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r, message=FALSE}
library(dplyr)
library(tidyr)

```

## Data Sources

Read in the data from KNB

```{r}
catch_original <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1")
```

## Clean and Reshape Data

* remove unnecessary columns
* check column types
* reshape data

ctrl + shft + m for the pipe

```{r}

catch_data <- catch_original %>% 
  mutate(Chinook = if_else(Chinook == "I", "1", Chinook)) %>% 
  mutate(Chinook = as.integer(Chinook)) %>% 
  select(-All, -notesRegCode)

```

To test which rows were causing problems in converting Chinook column to integer

```{r}
i = which(is.na(as.integer(catch_original$Chinook)))
catch_original[i,]
```
## Reshape data

```{r}

catch_long <- catch_data %>% 
  pivot_longer(cols = -c(Region, Year), names_to = "Species", values_to = "Catch")

catch_wide <- catch_long %>% 
  pivot_wider(names_from = Species, values_from = Catch)

```

## Group and Summarize

```{r}

mean_region <- catch_long %>% 
  group_by(Region) %>% 
  summarize(catch_mean = mean(Catch), .groups = "drop")

sum_year <- catch_long %>% group_by(Year) %>% 
  summarize(total_year = sum(Catch), num_obs = n(), .groups = "drop")

```

## Join to Regions Definitions

```{r}

region_defs <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1") %>% 
    select(code, mgmtArea) %>% 
    rename(Region = code)

```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = "Region")
```

Sam was here

Your presence was recorded ... 

checking if this works