---
title: "FunctionsPractice"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Functions Tutorial

## Load libraries

```{r}
library(ggplot2)

```


## Set up some test data

```{r create-data}
airtemps <- c(212, 78.3, 30, 32)

celcius1 <- (airtemps[1] - 32)*5/9
celcius2 <- (airtemps[2] - 32)*5/9
celcius3 <- (airtemps[3] - 32)*5/9
celcius4 <- (airtemps[4] - 32)*5/9

celcius1
celcius2
celcius3
celcius4
```
### Create a function to calculate celcius from farenheit

```{r create-function}

fahr_to_cel <- function(fahr) {
  celcius = (fahr - 32)*5/9
  return(celcius)
}

celcius = fahr_to_cel(airtemps)
celcius

```


### Create the opposite function

```{r inverse}

cel_to_fahr <- function(cel) {
  fahr = cel*9/5 + 32
  return(fahr)
}

fahrs = cel_to_fahr(celcius)
fahrs
```
### Return a dataframe with Farenheit, Celcius and Kelvin values

```{r return-dataframe}

convert_temps <- function(fahr) {
  celsius <- (fahr-32)*5/9
  kelvin <- celsius + 273.15
  return(list(fahr=fahr, celsius=celsius, kelvin=kelvin))
}

temps_df <- data.frame(convert_temps(seq(-100,100,10)))
temps_df

```

### Setting ggplot themes in general as a function (default value is 9)

can use *aes_string* or *formula* function to get around the string/symbol issue

```{r ggplot-theme}

custom_theme <- function(base_size = 9) {
    ggplot2::theme(
      axis.ticks       = ggplot2::element_blank(),
      text             = ggplot2::element_text(family = 'Helvetica', color = 'gray30', size = base_size),
      plot.title       = ggplot2::element_text(size = ggplot2::rel(1.25), hjust = 0.5, face = 'bold'),
      panel.background = ggplot2::element_blank(),
      legend.position  = 'right',
      panel.border     = ggplot2::element_blank(),
      panel.grid.minor = ggplot2::element_blank(),
      panel.grid.major = ggplot2::element_line(colour = 'grey90', size = .25),
      legend.key       = ggplot2::element_rect(colour = NA, fill = NA),
      axis.line        = ggplot2::element_blank()
      )
}

ggplot(temps_df, mapping=aes(x=fahr, y=celsius, color=kelvin)) +
    geom_point() +
    custom_theme(15)

```
### Mass producing scatterplots of same kind

```{r}

scatterplot <- function(df, point_size = 2, font_size=9) {
  ggplot(df, mapping=aes(x=fahr, y=celsius, color=kelvin)) +
    geom_point(size=point_size) +
    custom_theme(font_size)
}

scatterplot(temps_df, point_size=3, font_size = 16)

```

